---
title: "Simulation Study to Understand Sampling Distribution"
author: "Sulagna Barat | MDS202244"
subtitle: Problem 2
output:
  pdf_document: default
  html_document: default
---


$\underline{\textbf{Problem assignee}} \text{: Sulagna Baratl}$

$\underline{\textbf{Reviewer(s)}} \text{: Rohan Karthikeyan & Adarsha Mondal}$

## Problem 2 : Simulation Study to Understand Sampling Distribution

**Part A**
Suppose $X_1,X_2,\cdots,X_n\stackrel{iid}{\sim} Gamma(\alpha,\sigma)$, with pdf as
$$
f(x | \alpha,\sigma)=\frac{1}{\sigma^{\alpha}\Gamma(\alpha)}e^{- x/\sigma}x^{\alpha-1},~~~~0<x<\infty,
$$
The mean and variance are $E(X)=\alpha\sigma$ and $Var(X)=\alpha\sigma^2$. Note that `shape = ` $\alpha$ and `scale = ` $\sigma$.

1. Write a `function` in `R` which will compute the MLE of $\theta=\log(\alpha)$ using `optim` function in `R`. You can name it `MyMLE`
2. Choose `n=20`, and `alpha=1.5` and `sigma=2.2`
     (i) Simulate $\{X_1,X_2,\cdots,X_n\}$ from `rgamma(n=20,shape=1.5,scale=2.2)`
     (ii) Apply the `MyMLE` to estimate $\theta$ and append the value in a vector
     (iii) Repeat the step (i) and (ii) 1000 times
     (iv) Draw histogram of the estimated MLEs of $\theta$.
     (v) Draw a vertical line using `abline` function at the true value of $\theta$.
     (vi) Use `quantile` function on estimated $\theta$'s to find the 2.5 and 97.5-percentile points. 
3.  Choose `n=40`, and `alpha=1.5` and repeat the (2).
4.  Choose `n=100`, and `alpha=1.5` and repeat the (2).
5. Check if the gap between 2.5 and 97.5-percentile points are shrinking as sample size `n` is increasing?

*Hint*: Perhaps you should think of writing a single `function` where you will provide the values of `n`, `sim_size`, `alpha` and `sigma`; and it will return the desired output. 

```{r echo=FALSE}
library(tidyverse)
```

1. 
```{r warning=FALSE}
mle <- function(log_alpha, data, sigma) {
    l = sum(log(dgamma(data, shape = exp(log_alpha), scale = sigma)))
    # print(paste("l is ", l))
    return(-l)
}
MyMLE <- function(data, sigma) {
    log_alpha_initial <- log(mean(data)^2/var(data))
    # print(paste("log alpha initial is ", log_alpha_initial))
    estimator <- optim(log_alpha_initial,
                 mle,
                 data = data,
                 sigma = sigma)
    log_alpha_hat <- estimator$par
    return(log_alpha_hat)
}
```


```{r warning=FALSE}
get_estimates <- function(n, alpha, sigma) {
    estimates <- c()
    for (i in 1:1000) {
        samples <- rgamma(n, shape = alpha, scale = sigma)
        # print(paste("some of the samples are ", samples[1:5]))
        estimates <- append(estimates, MyMLE(data = samples, sigma = sigma))
    }
    return(estimates)
}
```

2. 
```{r warning=FALSE}
n = 20
alpha = 1.5
sigma = 2.2
estimated_mle <- tibble(get_estimates(n = n, alpha = alpha, sigma = sigma))
colnames(estimated_mle) <- c("estimate")
perc_2.5 <- quantile(estimated_mle$estimate, probs = 0.025, names = FALSE)
perc_97.5 <- quantile(estimated_mle$estimate, probs = 0.975, names = FALSE)
estimated_mle %>% 
    ggplot(aes(estimate)) +
    geom_histogram(color = "white", fill = "#5D5D5D") +
    geom_vline(xintercept = log(alpha), 
               size = 2,
               linetype = "dashed") +
    annotate("text", label = "Actual log(alpha)", x = 0.5, y = 95, color = "black") +
    geom_vline(xintercept = perc_2.5, 
               color = "#00B9FF", size = 1.5, linetype = "dashed") +
    annotate("text", label = "2.5 percentile", x = perc_2.5 + 0.1, y = 95, color = "#00B9FF") +
    geom_vline(xintercept = perc_97.5, 
               color = "#E08304", size = 1.5, linetype = "dashed") +
    annotate("text", label = "97.5 percentile", x = perc_97.5 + 0.1, y = 95, color = "#E08304") +
    labs(title = paste("n = ", n, ", alpha = ", alpha, ", sigma = ", sigma),
         x = "Estimated MLE",
         y = "Count")
diff_20 <- perc_97.5 - perc_2.5
```

3.
```{r warning=FALSE}
n = 40
alpha = 1.5
sigma = 2.2
estimated_mle <- tibble(get_estimates(n = n, alpha = alpha, sigma = sigma))
colnames(estimated_mle) <- c("estimate")
perc_2.5 <- quantile(estimated_mle$estimate, probs = 0.025, names = FALSE)
perc_97.5 <- quantile(estimated_mle$estimate, probs = 0.975, names = FALSE)
estimated_mle %>% 
    ggplot(aes(estimate)) +
    geom_histogram(color = "white", fill = "#5D5D5D") +
    geom_vline(xintercept = log(alpha), 
               size = 2,
               linetype = "dashed") +
    annotate("text", label = "Actual log(alpha)", x = log(alpha) + 0.1, y = 95, color = "black") +
    geom_vline(xintercept = perc_2.5, 
               color = "#00B9FF", size = 1.5, linetype = "dashed") +
    annotate("text", label = "2.5 percentile", x = perc_2.5 + 0.1, y = 95, color = "#00B9FF") +
    geom_vline(xintercept = perc_97.5, 
               color = "#E08304", size = 1.5, linetype = "dashed") +
    annotate("text", label = "97.5 percentile", x = perc_97.5 + 0.1, y = 95, color = "#E08304") +
    labs(title = paste("n = ", n, ", alpha = ", alpha, ", sigma = ", sigma),
         x = "Estimated MLE",
         y = "Count")
diff_40 <- perc_97.5 - perc_2.5
```

4. 
```{r warning=FALSE}
n = 100
alpha = 1.5
sigma = 2.2
estimated_mle <- tibble(get_estimates(n = n, alpha = alpha, sigma = sigma))
colnames(estimated_mle) <- c("estimate")
perc_2.5 <- quantile(estimated_mle$estimate, probs = 0.025, names = FALSE)
perc_97.5 <- quantile(estimated_mle$estimate, probs = 0.975, names = FALSE)
estimated_mle %>% 
    ggplot(aes(estimate)) +
    geom_histogram(color = "white", fill = "#5D5D5D") +
    geom_vline(xintercept = log(alpha), 
               size = 2,
               linetype = "dashed") +
    annotate("text", label = "Actual log(alpha)", x = log(alpha) + 0.05, y = 95, color = "black") +
    geom_vline(xintercept = perc_2.5, 
               color = "#00B9FF", size = 1.5, linetype = "dashed") +
    annotate("text", label = "2.5 percentile", x = perc_2.5 + 0.05, y = 95, color = "#00B9FF") +
    geom_vline(xintercept = perc_97.5, 
               color = "#E08304", size = 1.5, linetype = "dashed") +
    annotate("text", label = "97.5 percentile", x = perc_97.5 + 0.05, y = 95, color = "#E08304") +
    labs(title = paste("n = ", n, ", alpha = ", alpha, ", sigma = ", sigma),
         x = "Estimated MLE",
         y = "Count")
diff_100 <- perc_97.5 - perc_2.5
```
5.
```{r}
diff_20
diff_40
diff_100
```

We can see that the gap between the percentile points is decreasing as the sample size increases.

