---
title: "Computational Finance - Modelling Stock prices"
subtitle: Problem 5
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::hooks_markdown()$inline
library(ggplot2)
library(gridExtra)
library(reshape2)
library(ggpubr)
library(formatR)
library(quantmod)
```

Following piece of code download the prices of TCS since 2007

```{r warning=FALSE}
library(quantmod)
getSymbols('TCS.NS')
tail(TCS.NS)
```

Plot the adjusted close prices of TCS

```{r warning=FALSE}
print(nrow(TCS.NS$TCS.NS.Adjusted))
plot(TCS.NS$TCS.NS.Adjusted)
```

**Download the data of market index Nifty50**. The Nifty 50 index
indicates how the over all market has done over the similar period.

```{r warning=FALSE}
getSymbols('^NSEI')
tail(NSEI)
```

Plot the adjusted close value of Nifty50

```{r}
plot(NSEI$NSEI.Adjusted)
```

### Log-Return

We calculate the daily log-return, where log-return is defined as $$
r_t=\log(P_t)-\log(P_{t-1})=\Delta \log(P_t),
$$ where $P_t$ is the closing price of the stock on $t^{th}$ day.

```{r}
TCS_rt = diff(log(TCS.NS$TCS.NS.Adjusted))
Nifty_rt = diff(log(NSEI$NSEI.Adjusted))
retrn = cbind.xts(TCS_rt,Nifty_rt) 
retrn = na.omit(data.frame(retrn))

plot(retrn$NSEI.Adjusted,retrn$TCS.NS.Adjusted
     ,pch=20
     ,xlab='Market Return'
     ,ylab='TCS Return'
     ,xlim=c(-0.18,0.18)
     ,ylim=c(-0.18,0.18))
grid(col='grey',lty=1)
```

-   Consider the following model:

$$
r_{t}^{TCS}=\alpha + \beta r_{t}^{Nifty} + \varepsilon,
$$ where $\mathbb{E}(\varepsilon)=0$ and
$\mathbb{V}ar(\varepsilon)=\sigma^2$.

### Problem 1

> Estimate the parameters of the models $\theta=(\alpha,\beta,\sigma)$
> using the method of moments type plug-in estimator discussed in the
> class.

### Solution:

- #### Brief Introduction on Method of Moments

Method of Moments is a method of finding point estimators by equating
sample moments with the corresponding population moments.

We know, $j^{th}$ Moment of variable $X$,
$$M_{j}=\frac{1}{n}\sum_{i=1}^{n}X_{i}^{j}$$

- ##### Calculating Expectation by Method of Moments:

$$
\begin{align}
&\mathbb{E}(X)=M_{1}=m_{1}=\frac{1}{n}\sum_{i=1}^{n}X_{i}  \\
\end{align}
$$

After simplifying, we get `equation (1)` 
$$
\begin{align}
&\mathbb{E}(X)=M_{1}=\overline{X}
\end{align}
$$

- ##### Calculating Variance by Method of Moments:

$$
\begin{align}
&\mathbb{E}(X^{2})=M_{2}=m_{2}=\frac{1}{n}\sum_{i=1}^{n}X_{i}^{2} \\
or, &\mathbb{V}ar(X)+(\mathbb{E}(X))^{2}=M_{2}=\overline{X^{2}} \\
or, &\mathbb{V}ar(X)=M_{2}+(\mathbb{E}(X))^{2}=\overline{X^{2}}+(\mathbb{E}(X))^{2}
\end{align}
$$

After replacing from `equation (1)`, we get `equation (2)`,
$$\mathbb{V}ar(X)=M_{2}+M_{1}^{2}=\overline{X^{2}}+\overline{X}^{2}$$


#### Solving Approach

- ##### First Moments:

We have already seen (`equation (1)`) that the Expectation and First moment of a random variable is same.

```{r echo = FALSE}
r_nif_mean = mean(retrn$NSEI.Adjusted)
r_tcs_mean = mean(retrn$TCS.NS.Adjusted)
```

So, $\mathbb{E}(r^{Nifty})=$ `r r_nif_mean` and $\mathbb{E}(r^{TCS})=$
`r r_tcs_mean`

- ##### Second Moments:

```{r echo = FALSE}
r_nif_second_mom = mean(retrn$NSEI.Adjusted ** 2)
r_tcs_second_mom = mean(retrn$TCS.NS.Adjusted ** 2)
```

$\mathbb{E}(({r^{Nifty}})^2)=$ `r r_nif_second_mom` and
$\mathbb{E}(({r^{TCS}})^2)=$ `r r_tcs_second_mom`

- ##### Variance calculation:

```{r echo=FALSE}
r_nif_var = r_nif_second_mom + r_nif_mean ** 2
r_tcs_var = r_tcs_second_mom + r_tcs_mean ** 2
```

As we already have calculated First and Second moment for both the
variable, now Variance can be calculated using `equation (2)`.

$\mathbb{V}ar(r^{TCS}) = \mathbb{E}{(({r^{TCS}})^2)} + (\mathbb{E}(r^{TCS}))^{2}=$
`r r_tcs_var`

and

$\mathbb{V}ar(r^{Nifty}) = \mathbb{E}{(({r^{Nifty}})^2)} + (\mathbb{E}(r^{Nifty}))^{2}=$
`r r_nif_var`

- ##### Forming equations between parameters:

Applying property of Expectation on the given model, we get
    `equation (3)`

$$
\begin{align}
&\mathbb{E}(r^{TCS}) = \mathbb{E}(\alpha + \beta r^{Nifty} + \varepsilon) \\
\Rightarrow &\mathbb{E}(r^{TCS}) = \alpha + \beta*\mathbb{E}(r^{Nifty}) + \mathbb{E}(\varepsilon) \\
\Rightarrow &\mathbb{E}(r^{TCS}) = \alpha + \beta*\mathbb{E}(r^{Nifty})
\end{align}
$$

as, given $\mathbb{E}(\varepsilon) = 0$

Therefore putting values of $\mathbb{E}(r^{TCS})$ and
$\mathbb{E}(r^{Nifty})$ in `equation (3)` we get `equation (4)`

$\alpha +$ `r r_nif_mean` $\beta=$ `r r_tcs_mean`

Multiplying explanatory variable $r^{Nifty}$ on both sides of the
    model, we get,

$$
\begin{align}
r^{TCS} r^{Nifty} = \alpha r^{Nifty} + \beta (r^{Nifty})^{2} + \varepsilon r^{Nifty} \\
\end{align}
$$

Now, applying property of Expectation on the above equation
`equation (5)` is formed

$$
\begin{align}
&\mathbb{E}(r^{TCS} r^{Nifty}) = \mathbb{E}(\alpha r^{Nifty} + \beta (r^{Nifty})^{2} + \varepsilon r^{Nifty}) \\
\Rightarrow&\mathbb{E}(\alpha r^{Nifty}) + \mathbb{E}(\beta (r^{Nifty})^{2}) = \mathbb{E}(r^{TCS} r^{Nifty}) - \mathbb{E}(\varepsilon r^{Nifty}) \\
\Rightarrow &\alpha\mathbb{E}(r^{Nifty}) + \beta\mathbb{E}((r^{Nifty})^{2}) - \mathbb{E}(r^{TCS} r^{Nifty}) = \mathbb{E}(r^{TCS} r^{Nifty})
\end{align}
$$

```{r echo = FALSE}
tcs_nif_prod_mean = mean(retrn$TCS.NS.Adjusted * retrn$NSEI.Adjusted)
```

Replacing, expected values in the above equation we get, `equation (6)`

`r r_nif_mean` $\alpha+$ `r r_nif_second_mom` $\beta=$ `r tcs_nif_prod_mean`

Applying property of Variance on the given model, we get `equation (7)`

$$
\begin{align}
\mathbb{V}ar(r^{TCS}) &= \mathbb{V}ar(\alpha + \beta r^{Nifty} + \varepsilon) \\
\Rightarrow \mathbb{V}ar(r^{TCS})&=\mathbb{V}ar(\beta * r^{Nifty}) + \mathbb{V}ar(\varepsilon) \\
\Rightarrow \mathbb{V}ar(r^{TCS})&=\beta^{2} * \mathbb{V}ar(r^{Nifty}) + \sigma^{2} \\
\end{align}
$$

as, given $\mathbb{V}ar(\varepsilon) = \sigma^{2}$

Therefore putting values of $\mathbb{V}ar(r^{TCS})$ and
$\mathbb{V}ar(r^{Nifty})$ in `equation (7)`, we get `equation (8)`

`r r_nif_var` $\beta^{2}+\sigma^{2}=$ `r r_tcs_var`


#### Solving for values of the parameters

`equation (4)`

> $\alpha +$ `r r_nif_mean` $\beta=$ `r r_tcs_mean`

`equation (6)`

> `r r_nif_mean` $\alpha+$ `r r_nif_second_mom` $\beta=$ `r tcs_nif_prod_mean`

`equation (8)`

> `r r_nif_var` $\beta^{2}+\sigma^{2}=$ `r r_tcs_var`

- ##### Solving `equation (4)` & `(6)` we get the values of parameters $\alpha$ & $\beta$

```{r echo=FALSE}
A = rbind(c(1, r_nif_mean), c(r_nif_mean, r_nif_second_mom))
B = c(r_tcs_mean, tcs_nif_prod_mean)
values = solve(A,B)
alpha_mm = values[1]; beta_mm = values[2]
sigma_mm = (r_tcs_var - (r_nif_var*(beta_mm**2)))**0.5
```

- ##### Now putting the value of $\beta$ into `equation (8)` we get the value of $\sigma$


### Problem 2

> Estimate the parameters using the `lm` built-in function of `R`. Note
> that `lm` using the OLS method.

### Solution:

### Problem 3

> Fill-up the following table

| Parameters | Method of Moments | OLS |
|------------|-------------------|-----|
| $\alpha$   |  `r alpha_mm`     |     |
| $\beta$    |  `r beta_mm`      |     |
| $\sigma$   |  `r sigma_mm`     |     |

### Problem 4

> If the current value of Nifty is 18000 and it goes up to 18200. The
> current value of TCS is Rs. 3200/-. How much you can expect TCS price
> to go up?

### Solution:
