---
title: "Simulation Study to Understand Sampling Distribution"
author: "Sulagna Barat | MDS202244"
subtitle: Problem 2
output:
  pdf_document: default
  html_document: default
---


$\underline{\textbf{Problem assignee}} \text{: Sulagna Barat}$

$\underline{\textbf{Reviewer(s)}} \text{: Rohan Karthikeyan and Adarsha Mondal}$

## Problem 2 : Simulation Study to Understand Sampling Distribution

**Part A**
Suppose $X_1,X_2,\cdots,X_n\stackrel{iid}{\sim} Gamma(\alpha,\sigma)$, with pdf as
$$
f(x | \alpha,\sigma)=\frac{1}{\sigma^{\alpha}\Gamma(\alpha)}e^{- x/\sigma}x^{\alpha-1},~~~~0<x<\infty,
$$
The mean and variance are $E(X)=\alpha\sigma$ and $Var(X)=\alpha\sigma^2$. Note that `shape = ` $\alpha$ and `scale = ` $\sigma$.

1. Write a `function` in `R` which will compute the MLE of $\theta=\log(\alpha)$ using `optim` function in `R`. You can name it `MyMLE`
2. Choose `n=20`, and `alpha=1.5` and `sigma=2.2`
     (i) Simulate $\{X_1,X_2,\cdots,X_n\}$ from `rgamma(n=20,shape=1.5,scale=2.2)`
     (ii) Apply the `MyMLE` to estimate $\theta$ and append the value in a vector
     (iii) Repeat the step (i) and (ii) 1000 times
     (iv) Draw histogram of the estimated MLEs of $\theta$.
     (v) Draw a vertical line using `abline` function at the true value of $\theta$.
     (vi) Use `quantile` function on estimated $\theta$'s to find the 2.5 and 97.5-percentile points. 
3.  Choose `n=40`, and `alpha=1.5` and repeat the (2).
4.  Choose `n=100`, and `alpha=1.5` and repeat the (2).
5. Check if the gap between 2.5 and 97.5-percentile points are shrinking as sample size `n` is increasing?

*Hint*: Perhaps you should think of writing a single `function` where you will provide the values of `n`, `sim_size`, `alpha` and `sigma`; and it will return the desired output. 

```{r echo=FALSE, messages=FALSE}
library(tidyverse)
```
\newpage
Let $x_1,x_2,\cdots,x_n$ be a random sample of a popuation with pdf $f(x;\theta)$, where $\theta$ is a parameter. Consider a function, \[f(x_1,x_2,...,x_n;\theta)=\prod_{i=1}^n f(x_i;\theta)\]
When $x_1,x_2,\cdots,x_n$ are given, then $f(x_1,x_2,...,x_n;\theta)$ is a function of $\theta$ only, that we call likelihood function of $\theta$, denoted by $L(\theta)$. An estimate of $\theta$ for which $L(\theta)$ is maximum (consequently $log L(\theta)$ is also maximum) is suggested by the 'Maximum Likelihood'.
Maximum Likelihood Estimator of a parameter $\theta$ is a consistent estimator of $\theta$.
\newline 
Here we are using $optim$ function to compute MLE of $\theta=\log(\alpha)$ and naming it as MyMLE.\newline
1. 
```{r warning=FALSE}
mle <- function(log_alpha, data, sigma) {
    l = sum(log(dgamma(data, shape = exp(log_alpha), scale = sigma)))
    return(-l)
}
MyMLE <- function(data, sigma) {
    log_alpha_init <- log(mean(data)^2/var(data))
    estimator <- optim(log_alpha_init,
                 mle,
                 data = data,
                 sigma = sigma)
    log_alpha_cap <- estimator$par
    return(log_alpha_cap)
}
```

```{r warning=FALSE}
get_estim <- function(n, alpha, sigma) {
    estim <- c()
    for (i in 1:1000) {
        samples <- rgamma(n, shape = alpha, scale = sigma)
        estim <- append(estim, MyMLE(data = samples, sigma = sigma))
    }
    return(estim)
}
```
\newpage
2. 
```{r warning=FALSE}
n = 20
alpha = 1.5
sigma = 2.2
estim_mle <- tibble(get_estim(n = n, alpha = alpha, sigma = sigma))
colnames(estim_mle) <- c("estim")
percentile_2.5 <- quantile(estim_mle$estim, probs = 0.025, names = FALSE)
percentile_97.5 <- quantile(estim_mle$estim, probs = 0.975, names = FALSE)
estim_mle %>% 
    ggplot(aes(estim)) +
    geom_histogram(color = "red", fill = "yellow") +
    geom_vline(xintercept = log(alpha), 
               size = 1,
               linetype = "solid") +
    annotate("text", label = "log(alpha) value", x = 0.5, y = 95, color = "black") +
    geom_vline(xintercept = percentile_2.5, 
               color = "dark red", size = 1, linetype = "dotted") +
    annotate("text", label = "2.5 percentile", x = percentile_2.5 + 0.1, y = 95, color = "black") +
    geom_vline(xintercept = percentile_97.5, 
               color = "dark green", size = 1, linetype = "dotted") +
    annotate("text", label = "97.5 percentile", x = percentile_97.5 + 0.1, y = 95, color = "black") +
    labs(title = paste("n = ", n, ", alpha = ", alpha, ", sigma = ", sigma),
         x = "Estimated MLE",
         y = "Count")
diff_20 <- percentile_97.5 - percentile_2.5
```
\newpage
3.
```{r warning=FALSE}
n = 40
alpha = 1.5
sigma = 2.2
estim_mle <- tibble(get_estim(n = n, alpha = alpha, sigma = sigma))
colnames(estim_mle) <- c("estim")
percentile_2.5 <- quantile(estim_mle$estim, probs = 0.025, names = FALSE)
percentile_97.5 <- quantile(estim_mle$estim, probs = 0.975, names = FALSE)
estim_mle %>% 
    ggplot(aes(estim)) +
    geom_histogram(color = "red", fill = "yellow") +
    geom_vline(xintercept = log(alpha), 
               size = 1,
               linetype = "solid") +
    annotate("text", label = "log(alpha)", x = log(alpha) + 0.1, y = 95, color = "black") +
    geom_vline(xintercept = percentile_2.5, 
               color = "dark red", size = 1, linetype = "dotted") +
    annotate("text", label = "2.5 percentile", x = percentile_2.5 + 0.1, y = 95, color = "black") +
    geom_vline(xintercept = percentile_97.5, 
               color = "dark green", size = 1, linetype = "dotted") +
    annotate("text", label = "97.5 percentile", x = percentile_97.5 + 0.1, y = 95, color = "black") +
    labs(title = paste("n = ", n, ", alpha = ", alpha, ", sigma = ", sigma),
         x = "Estimated MLE",
         y = "Count")
diff_40 <- percentile_97.5 - percentile_2.5
```
\newpage
4. 
```{r warning=FALSE}
n = 100
alpha = 1.5
sigma = 2.2
estim_mle <- tibble(get_estim(n = n, alpha = alpha, sigma = sigma))
colnames(estim_mle) <- c("estim")
percentile_2.5 <- quantile(estim_mle$estim, probs = 0.025, names = FALSE)
percentile_97.5 <- quantile(estim_mle$estim, probs = 0.975, names = FALSE)
estim_mle %>% 
    ggplot(aes(estim)) +
    geom_histogram(color = "red", fill = "yellow") +
    geom_vline(xintercept = log(alpha), 
               size = 1,
               linetype = "solid") +
    annotate("text", label = "log(alpha)", x = log(alpha) + 0.05, y = 95, color = "black") +
    geom_vline(xintercept = percentile_2.5, 
               color = "dark red", size = 1, linetype = "dotted") +
    annotate("text", label = "2.5 percentile", x = percentile_2.5 + 0.05, y = 95, color = "black") +
    geom_vline(xintercept = percentile_97.5, 
               color = "dark green", size = 1, linetype = "dotted") +
    annotate("text", label = "97.5 percentile", x = percentile_97.5 + 0.05, y = 95, color = "black") +
    labs(title = paste("n = ", n, ", alpha = ", alpha, ", sigma = ", sigma),
         x = "Estimated MLE",
         y = "Count")
diff_100 <- percentile_97.5 - percentile_2.5
```
\newpage
5.
```{r}
diff_20
diff_40
diff_100
```
Conclusion:
Clearly, the gap between the percentile points is decreasing as the sample size increases.

